How General Health Index (GHI) performs better to predict health levels

### 2024-11-30 paper/paper.qmd
maybe working model

@@

>>>>

library(tidyverse)
library(janitor)
library(arrow)
library(dplyr)
```

@@

## BMI Classification:
```{r}
#| warning: false
#| message: false
bmi_canada_gov <- data.frame(
  Category = c("Underweight", "Normal weight", "Overweight", "Obese - Class I", "Obese - Class II", "Obese - Class III"),
  `BMI Range` = c("Below 18.5", "18.5–24.9", "25.0–29.9", "30.0–34.9", "35.0–39.9", "40.0 and above")
)

# Create a table using knitr::kable
knitr::kable(
  bmi_canada_gov,
  col.names = c("Category", "BMI Range"),
  caption = "BMI Categories as per the Government of Canada Guidelines"
)
```
### The difference between an underweight and normal person:
Let us take a closer look at the BMI levels below 25 for underweight and normal weight adults.

```{r}
#| warning: false
#| message: false
body_mass_data_raw <- read_parquet("../data/raw_data/body_mass_data.parquet")
body_mass_clean <- body_mass_data_raw %>%
  clean_names() %>%
  select(gender_m1f2, age, height_cm, w_kg, bmi, fat_mass_kg, fm, ffm_kg, bone_mass_kg, muscle_mass_kg) %>%
  mutate(
    bmi_category = case_when(
      bmi < 18.5 ~ "Underweight",
      bmi >= 18.5 & bmi <= 24.9 ~ "Normal weight",
      bmi >= 25.0 & bmi <= 29.9 ~ "Overweight",
      bmi >= 30.0 & bmi <= 34.9 ~ "Obese - Class I",
      bmi >= 35.0 & bmi <= 39.9 ~ "Obese - Class II",
      bmi >= 40.0 ~ "Obese - Class III"
    ))

# Data Section: Part 1: What is the difference between an underweight / athletic person / normal person:
bmi_under_25 <- body_mass_clean %>%
  filter(bmi < 25)

ggplot(bmi_under_25, aes(x = bmi, y = fm)) +
  geom_point(aes(color = bmi_category), alpha = 0.7) +
  labs(
    title = "BMI vs. Fat Mass by BMI Category",
    x = "BMI",
    y = "Fat Mass (kg)",
    color = "BMI Category"
  ) +
  theme_minimal()

### The difference between a normal and overweight person:
```{r}
bmi_normal_over <- body_mass_clean %>%
  filter(18.5 < bmi & bmi < 30)

# this is what i chose
ggplot(bmi_normal_over, aes(x = bmi, y = fm)) +
  geom_point(aes(color = bmi_category), alpha = 0.7) +
  labs(
    title = "BMI vs. Fat Mass by BMI Category",
    x = "BMI",
    y = "Fat Mass (kg)",
    color = "BMI Category"
  ) +
  theme_minimal()
```

### All categories
```{r}
# this is what i chose
ggplot(body_mass_clean, aes(x = bmi, y = fm)) +
  geom_point(aes(color = bmi_category), alpha = 0.7) +
  geom_vline(xintercept = c(18.5, 25, 30), linetype = "dashed", color = "black", linewidth = 0.5, alpha = 0.5) +
  labs(
    title = "BMI vs. Fat Mass by BMI Category",
    x = "BMI",
    y = "Fat Mass (kg)",
    color = "BMI Category"
  ) +
  theme_minimal()
```
